Project Report\\\textsc{DEVS Ex Machina}

### 2015-05-25 manual/projectrapport.tex
Updated performance stats for parallel with zombie round fix

\chapter{Logger analysis}

>>>>

\chapter{Performance analysis}
To measure the performance of \textsc{DEVS Ex Machina} under more demanding conditions, \textbf{DEVStone} and \textbf{PHOLD} benchmarks were used in combination with the \texttt{perf} profiler tool. Every time we compare the performance of our implementation with \textit{PythonPDEVS}.\\
Note: at the moment, no major research has gone into the use of compiler optimizations yet. Potentially, there is still speed to be gained here.\\
\\
{\footnotesize[All the following tests were performed using an \textit{Intel Core i5 M520} CPU at 2.40GHz, parallel testing was done using 2 simulation cores.]}

\section{Classic DEVS}
Using Classic DEVS we see noticeable speedups compared to PyPDEVS. Both tests are completed in far less time, using the CPU more efficiently.

\subsection{DEVStone}
The DEVStone benchmark used for both implementations has a width of 2 and a depth of 3.
\paragraph{\textsc{DEVS Ex Machina}}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_devstone classic 2 3

 Performance counter stats for './dxexmachina_devstone classic 2 3':

        174.007707 task-clock (msec)         #    1.111 CPUs utilized          
               184 context-switches          #    0.001 M/sec                  
               212 cpu-migrations            #    0.001 M/sec                  
               696 page-faults               #    0.004 M/sec                  
       463,684,386 cycles                    #    2.665 GHz                     [83.13%]
       196,709,921 stalled-cycles-frontend   #   42.42% frontend cycles idle    [85.77%]
       103,770,146 stalled-cycles-backend    #   22.38% backend  cycles idle    [68.62%]
       557,355,313 instructions              #    1.20  insns per cycle        
                                             #    0.35  stalled cycles per insn [86.11%]
       131,411,927 branches                  #  755.208 M/sec                   [85.62%]
         1,720,031 branch-misses             #    1.31% of all branches         [83.38%]

       0.156664362 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{PythonPDEVS}}
\begin{Verbatim}[fontsize=\small]
$ perf stat python experiment.py default 2 false > out.txt

 Performance counter stats for 'python experiment.py default 2 false':

        709.776558 task-clock (msec)         #    0.989 CPUs utilized          
               191 context-switches          #    0.269 K/sec                  
                34 cpu-migrations            #    0.048 K/sec                  
             2,942 page-faults               #    0.004 M/sec                  
     1,687,298,167 cycles                    #    2.377 GHz                     [82.71%]
       866,759,527 stalled-cycles-frontend   #   51.37% frontend cycles idle    [83.25%]
       562,247,464 stalled-cycles-backend    #   33.32% backend  cycles idle    [67.62%]
     1,672,919,036 instructions              #    0.99  insns per cycle        
                                             #    0.52  stalled cycles per insn [83.82%]
       375,443,818 branches                  #  528.961 M/sec                   [83.25%]
        14,867,814 branch-misses             #    3.96% of all branches         [83.28%]

       0.717576473 seconds time elapsed
\end{Verbatim}
\subsection{PHOLD}
\paragraph{\textsc{DEVS Ex Machina}}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_phold classic 1 10 0 10

 Performance counter stats for './dxexmachina_phold classic 1 10 0 10':

         15.847971 task-clock (msec)         #    1.047 CPUs utilized          
                15 context-switches          #    0.946 K/sec                  
                28 cpu-migrations            #    0.002 M/sec                  
               710 page-faults               #    0.045 M/sec                  
        27,365,192 cycles                    #    1.727 GHz                     [74.72%]
        12,605,461 stalled-cycles-frontend   #   46.06% frontend cycles idle    [73.78%]
         7,424,569 stalled-cycles-backend    #   27.13% backend  cycles idle    [86.04%]
        31,095,649 instructions              #    1.14  insns per cycle        
                                             #    0.41  stalled cycles per insn
         6,955,945 branches                  #  438.917 M/sec                  
           132,996 branch-misses             #    1.91% of all branches         [74.99%]

       0.015131930 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{PythonPDEVS}}
\begin{Verbatim}[fontsize=\small]
$ perf stat python experiment.py default 2 false > out.txt

 Performance counter stats for 'python experiment.py default 2 false':

        709.776558 task-clock (msec)         #    0.989 CPUs utilized          
               191 context-switches          #    0.269 K/sec                  
                34 cpu-migrations            #    0.048 K/sec                  
             2,942 page-faults               #    0.004 M/sec                  
     1,687,298,167 cycles                    #    2.377 GHz                     [82.71%]
       866,759,527 stalled-cycles-frontend   #   51.37% frontend cycles idle    [83.25%]
       562,247,464 stalled-cycles-backend    #   33.32% backend  cycles idle    [67.62%]
     1,672,919,036 instructions              #    0.99  insns per cycle        
                                             #    0.52  stalled cycles per insn [83.82%]
       375,443,818 branches                  #  528.961 M/sec                   [83.25%]
        14,867,814 branch-misses             #    3.96% of all branches         [83.28%]

       0.717576473 seconds time elapsed
\end{Verbatim}

\section{Parallel DEVS}
Also in parallel simulations \textsc{DEVS Ex Machina} completes quite a bit quicker, again achieving higher CPU utilization.
Additionally we can see that Conservative PDEVS performs consistently significantly faster than Parallel PDEVS, in the case of PHOLD.

\subsection{DEVStone}
\paragraph{\textsc{DEVS Ex Machina} - Optimistic}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_devstone opdevs 2 2 3

 Performance counter stats for './dxexmachina_devstone opdevs 2 2 3':

         95.265846 task-clock (msec)         #    0.499 CPUs utilized          
               191 context-switches          #    0.002 M/sec                  
                15 cpu-migrations            #    0.157 K/sec                  
             1,229 page-faults               #    0.013 M/sec                  
       197,702,110 cycles                    #    2.075 GHz                     [82.53%]
        95,414,988 stalled-cycles-frontend   #   48.26% frontend cycles idle    [83.54%]
        47,283,951 stalled-cycles-backend    #   23.92% backend  cycles idle    [76.71%]
       199,327,993 instructions              #    1.01  insns per cycle        
                                             #    0.48  stalled cycles per insn [91.61%]
        47,887,986 branches                  #  502.677 M/sec                   [86.14%]
           931,688 branch-misses             #    1.95% of all branches         [75.36%]

       0.191066656 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{DEVS Ex Machina} - Conservative}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_devstone cpdevs 2 2 3

 Performance counter stats for './dxexmachina_devstone cpdevs 2 2 3':

        129.292299 task-clock (msec)         #    0.555 CPUs utilized          
               672 context-switches          #    0.005 M/sec                  
                15 cpu-migrations            #    0.116 K/sec                  
             2,177 page-faults               #    0.017 M/sec                  
       308,983,665 cycles                    #    2.390 GHz                     [74.75%]
       151,616,998 stalled-cycles-frontend   #   49.07% frontend cycles idle    [84.43%]
        82,891,133 stalled-cycles-backend    #   26.83% backend  cycles idle    [76.81%]
       276,296,034 instructions              #    0.89  insns per cycle        
                                             #    0.55  stalled cycles per insn [88.68%]
        66,513,459 branches                  #  514.443 M/sec                   [87.86%]
         1,390,075 branch-misses             #    2.09% of all branches         [77.49%]

       0.233046274 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{PythonPDEVS}}
\begin{Verbatim}[fontsize=\small]
$ perf stat mpirun -np 2 python experiment.py default 2 false > out.txt

 Performance counter stats for 'mpirun -np 2 python experiment.py default 2 false':

        787.125992 task-clock (msec)         #    0.371 CPUs utilized          
             1,120 context-switches          #    0.001 M/sec                  
                91 cpu-migrations            #    0.116 K/sec                  
             8,684 page-faults               #    0.011 M/sec                  
     1,786,217,349 cycles                    #    2.269 GHz                     [83.85%]
       935,480,076 stalled-cycles-frontend   #   52.37% frontend cycles idle    [84.18%]
       619,366,137 stalled-cycles-backend    #   34.67% backend  cycles idle    [65.26%]
     1,796,463,414 instructions              #    1.01  insns per cycle        
                                             #    0.52  stalled cycles per insn [84.16%]
       403,042,819 branches                  #  512.044 M/sec                   [84.70%]
        13,210,328 branch-misses             #    3.28% of all branches         [82.68%]

       2.123888919 seconds time elapsed
\end{Verbatim}

\subsection{PHOLD}
\paragraph{\textsc{DEVS Ex Machina} - Optimistic}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_phold opdevs 2 1 10 0 10

 Performance counter stats for './dxexmachina_phold opdevs 2 1 10 0 10':

         20.741980 task-clock (msec)         #    0.092 CPUs utilized          
                73 context-switches          #    0.004 M/sec                  
                 8 cpu-migrations            #    0.386 K/sec                  
               760 page-faults               #    0.037 M/sec                  
        28,281,718 cycles                    #    1.364 GHz                     [42.03%]
        14,899,039 stalled-cycles-frontend   #   52.68% frontend cycles idle    [79.62%]
         8,373,668 stalled-cycles-backend    #   29.61% backend  cycles idle   
        29,601,476 instructions              #    1.05  insns per cycle        
                                             #    0.50  stalled cycles per insn
         6,754,172 branches                  #  325.628 M/sec                  
           135,220 branch-misses             #    2.00% of all branches         [81.22%]

       0.225834153 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{DEVS Ex Machina} - Conservative}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_phold cpdevs 2 1 10 0 10

 Performance counter stats for './dxexmachina_phold cpdevs 2 1 10 0 10':

         28.509489 task-clock (msec)         #    0.286 CPUs utilized          
                95 context-switches          #    0.003 M/sec                  
                11 cpu-migrations            #    0.386 K/sec                  
               813 page-faults               #    0.029 M/sec                  
        54,757,291 cycles                    #    1.921 GHz                     [63.45%]
        29,057,227 stalled-cycles-frontend   #   53.07% frontend cycles idle    [86.87%]
        15,602,463 stalled-cycles-backend    #   28.49% backend  cycles idle    [90.08%]
        53,326,834 instructions              #    0.97  insns per cycle        
                                             #    0.54  stalled cycles per insn
        12,449,801 branches                  #  436.690 M/sec                  
           219,155 branch-misses             #    1.76% of all branches         [64.68%]

       0.099591165 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{PythonPDEVS}}
PHOLD in parallel on PyPDEVS crashed immediately with an MPI error, hence a comparison was not possible for now.

\section{Compared to \textsc{ADEVS}}
Since the DEVStone benchmark for \textsc{ADEVS} does not support different depths, a test with greater width was performed instead. As we can see, \textsc{ADEVS} is still quite a bit faster.
\paragraph{\textsc{DEVS Ex Machina} - Classic DEVS}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./dxexmachina_devstone classic 32 0

 Performance counter stats for './dxexmachina_devstone classic 32 0':

        800.259000 task-clock (msec)         #    1.081 CPUs utilized          
             2,178 context-switches          #    0.003 M/sec                  
               217 cpu-migrations            #    0.271 K/sec                  
               767 page-faults               #    0.958 K/sec                  
     2,199,059,849 cycles                    #    2.748 GHz                     [83.12%]
       862,325,341 stalled-cycles-frontend   #   39.21% frontend cycles idle    [83.36%]
       432,658,584 stalled-cycles-backend    #   19.67% backend  cycles idle    [67.64%]
     2,794,549,736 instructions              #    1.27  insns per cycle        
                                             #    0.31  stalled cycles per insn [84.75%]
       661,425,743 branches                  #  826.515 M/sec                   [84.04%]
         9,605,392 branch-misses             #    1.45% of all branches         [82.83%]

       0.739954252 seconds time elapsed
\end{Verbatim}
\paragraph{\textsc{ADEVS} - Classic DEVS}
\begin{Verbatim}[fontsize=\small]
$ perf stat ./sim 32 0

 Performance counter stats for './sim 32 0':

         15.465233 task-clock (msec)         #    0.976 CPUs utilized          
                 1 context-switches          #    0.065 K/sec                  
                 1 cpu-migrations            #    0.065 K/sec                  
               340 page-faults               #    0.022 M/sec                  
        22,395,954 cycles                    #    1.448 GHz                     [74.20%]
         6,618,757 stalled-cycles-frontend   #   29.55% frontend cycles idle    [74.17%]
         2,895,009 stalled-cycles-backend    #   12.93% backend  cycles idle    [87.88%]
        31,797,999 instructions              #    1.42  insns per cycle        
                                             #    0.21  stalled cycles per insn
         6,645,494 branches                  #  429.705 M/sec                  
           250,467 branch-misses             #    3.77% of all branches         [74.03%]

       0.015852985 seconds time elapsed
\end{Verbatim}
